Total Score

### 2023-02-22 src/MSAtotal_scores.Rmd
trying to create a function to calculate weighted scores

criteria_all <- criteria_all[, col_order]

@@

```{r original weights, eval = FALSE}
# weight criteria for Owen
criteria_weights_owen <- c(0.37, 0.12, 0.22, 0.07, 0.0, 0.07, 0.15)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(total_score_owen = cbsa, #this is just a place holder
         real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)
  )

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i])
  criteria_all$total_score_owen[i] = weighted.mean(cbsa_data, criteria_weights_owen)
}

@@

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(final_tot_students = cbsa,
         stud_EAL_total = cbsa,
         real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)) #this is just a place holder

@@

```{r student weights with EAL, eval = FALSE}
size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$inverted_n_eal_score[i]
                  )
  criteria_all$stud_EAL_total[i] = weighted.mean(cbsa_data, criteria_weights_students)
}
```

```{r OWEN FINAL WEIGHTS climate risk 50 and proportional weights for others}

weights_climate50_prop <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(owen_FINAL_score = cbsa, #this is just a place holder
         score_climate50_prop_EAL = cbsa)

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i]
                  )
  criteria_all$owen_FINAL_score[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
}

@@

```{r Climate with 50 & proportional others with EAL, eval=FALSE}
### Using climate weight 50 and proportional others with EAL
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$inverted_n_eal_score[i]
                  )
  criteria_all$score_climate50_prop_EAL[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
}

```{r climate risk 35 and Jan 31 weights from owen, eval = FALSE}

weights_climate35_owen <- c(0.25, 0.1, 0.15, 0.075, 0.0, 0.075, 0.35)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(score_climate35_owen = cbsa)#this is just a place holder

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i]
                  )
  criteria_all$score_climate35_owen[i] = weighted.mean(cbsa_data, weights_climate35_owen) 
}
```

@@

criteria_all <- criteria_all %>% 
  rename(financial_score = normalized_irr,
         co2_e = inverted_n_total_co2,
         health_impact = inverted_n_health_emissions_cost,
         electricity = normalized_annual_renewable,
         climate_risk = normalized_inverted_climate_risk,
         climate_EAL_risk = inverted_n_eal_score) %>% 
  mutate_if(is.numeric, ~round(., 5))

@@

@@

```{r}
## read in weighted criteria sheet
student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud") %>% 
  unite("msa_state", city_msa:state, sep = ", ")

>>>>

criteria_all <- criteria_all[, col_order] %>% 
  select(!inverted_n_eal_score)

@@

```{r}
# weighting criteria
weighted_criteria <- criteria_all %>% 
  select(4:10)

weighted <- t(t(weighted_criteria)*criteria_weights_client)
```

@@

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(final_tot_students = cbsa,
         real_estate_score = as.numeric(real_estate_score),
         landlord_score = as.numeric(landlord_score)) #this is just a place holder

@@

```{r Client FINAL WEIGHTS}

client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)

#make landlord_score column
criteria_all <- criteria_all %>% 
  mutate(client_FINAL_score = cbsa, #this is just a place holder
         score_climate50_prop_EAL = cbsa)

size <- dim(criteria_all)
for (i in 1:size[1]) {
  cbsa_data <- c(criteria_all$real_estate_score[i],
                criteria_all$landlord_score[i],
                 criteria_all$normalized_irr[i],
                 criteria_all$inverted_n_total_co2[i],
                criteria_all$inverted_n_health_emissions_cost[i],
                criteria_all$normalized_annual_renewable[i],
                criteria_all$normalized_inverted_climate_risk[i]
                  )
  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
}

@@

@@

criteria_all <- criteria_all %>% 
  rename(financial_score = normalized_irr,
         co2_e = inverted_n_total_co2,
         health_impact = inverted_n_health_emissions_cost,
         electricity = normalized_annual_renewable,
         climate_risk = normalized_inverted_climate_risk) %>% 
  mutate_if(is.numeric, ~round(., 5))

@@

### Data Visualization- barcharts

@@

```{r student data viz}
## read in weighted criteria sheet
student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud") %>% 
  unite("msa_state", city_msa:state, sep = ", ")
